# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data


```{r loadingData}
library(lubridate)
library(lattice)
activity <- read.csv("activity.csv")
```

"...This device collects data at 5 minute intervals through out the day..."
So, we should have 288 intervals per day (288 minutes x 5 = 1440 minutes = 24 hours) and the minimum and maximum value should be 0 and 1435 respectively. Each day has 288 intervals, but the maximum value is not 1435.

```{r intervalRange}
rangeInterval <- range(activity$interval)
```
The range is `r rangeInterval` which it doesnt make sense for our purposes. Lets plot it:

```{r wrongInterval}
plot(1:288, as.numeric(levels(factor(activity$interval))), type="l", xlab = "number of intervals", ylab = "minutes")
```

There are regular gaps between several pairs of intervals. Lets fix them with new labels from 0 to 1435:

```{r fixingIntervals}
activity$interval <- factor(activity$interval, labels=seq(0, by=5, length=288))
plot(1:288, as.numeric(levels(activity$interval)), type="l", xlab = "number of intervals", ylab = "minutes")
```

## What is mean total number of steps taken per day?


```{r computeMean}
stepsPerDay <- with(activity, tapply(steps, date, sum))
meanStepsPerDay <- mean(stepsPerDay, na.rm=T)
```
The mean total number of steps taken per day is `r meanStepsPerDay`.

```{r computeMedian}
medianStepsPerDay <- median(stepsPerDay, na.rm=T)
```
The total median is `r medianStepsPerDay`.

The histogram looks like this:

```{r newHistogram, fig.height=4}
hist(stepsPerDay, col="blue", breaks=10, xlab = "Steps per day")
```

## What is the average daily activity pattern?

```{r activityPattern, fig.height=4}
stepsPerInterval <- with(activity, tapply(steps, interval, mean, na.rm=TRUE))
plot(levels(activity$interval), stepsPerInterval, type="l", xlab = "Minutes", ylab ="Steps per interval")
``` 

```{r maxStepsPerInterval}
maxStepsPerInterval <- max(stepsPerInterval)
intervalOfmaxSteps <- names(stepsPerInterval[stepsPerInterval == maxStepsPerInterval])
timeOfMaxInterval <- format(Sys.Date() + minutes(intervalOfmaxSteps), "%H:%M")
```
The maximum steps per interval is `r maxStepsPerInterval` , which corresponds to the interval `r intervalOfmaxSteps`, at `r timeOfMaxInterval`.


## Imputing missing values

I think it is better to fill NAs with the mean of the interval, that we calculated before:

```{r missingValues, fig.height=4}
noNAsActivity <- activity
noNAsActivity$steps <- apply(noNAsActivity, 1, 
	function (x) 
		if (is.na(x["steps"])) 
			as.numeric(stepsPerInterval[x["interval"]]) 
		else 
			as.numeric(x["steps"]))
head(noNAsActivity)
noNAsStepsPerDay <- with(noNAsActivity, tapply(steps, date, sum))
hist(noNAsStepsPerDay, col="green", breaks=10, xlab = "Steps per day")
```

```{r noNAsMean}
noNAsMean <- mean(noNAsStepsPerDay)
```

The mean is `r noNAsMean`, same as before because we have used the average intervals for fill the NAs





## Are there differences in activity patterns between weekdays and weekends?

```{r weekdyasPatterns}
noNAsActivity$weekday <- wday(ymd(noNAsActivity$date), label=TRUE)
noNAsActivity$week <- sapply(noNAsActivity$weekday,
	function (x)
		if (x %in% c("Sat","Sun"))
			"weekend"
		else
			"weekday"
	)
stepsPerIntervalWeekDay <- with(noNAsActivity[noNAsActivity$week == "weekday",], 
	tapply(steps, interval, mean))
stepsPerIntervalWeekEnd <- with(noNAsActivity[noNAsActivity$week == "weekend",], 
	tapply(steps, interval, mean))

stepsPerInterval <- data.frame(interval = rep(seq(0, by=5, length=288), 2),
	steps = c(stepsPerIntervalWeekDay, stepsPerIntervalWeekEnd), 
	week = c(rep("weekday", 288), rep("weekend", 288)))

xyplot(steps~interval | week, data = stepsPerInterval, layout = c(1,2), type="l")
```


So... yes, the are differences between weedays and weekends. He/She looks more active in weekends :)

